---
title: "Writing Research Papers"
author: |
  | Author 1^[Corresponding author: email@email.com] $^1$, Author 2 $^1$, Author 3 $^2$
  | $^1$Affiliation1,  $^2$Affiliation2

abstract: |
  This is a guideline for students writing in biological science, and other forms of science writing.
  
bibliography: bib/biosci.bib       #Path to your .bib file
bibli0-style: apalike
#csl: Path to your .csl file #sets the style of your bibliography
output:
  bookdown::pdf_document2:
toc: no
fig_caption: yes
always_allow_html: yes
# These are LaTex settings to take care of floating figures/tables, line spacing, etc
header-includes:
  - \usepackage{endfloat}    
  - \usepackage{setspace}\doublespacing
 #- \usepackage{lineno}
 # - \linenumbers
---

 ```{r Setup, message=FALSE, warning=FALSE, include=FALSE}
# knitr::opts_chunk$set(fig.pos = 'p')  # Places figures on their own pages
knitr::opts_chunk$set(out.width = '100%', dpi=300)

# I usually load my libraries up front to keep things organized
library(bookdown)
library(tidyverse)
#library(kableExtra)
#library(ggthemes)
#library(ggExtra)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,           # show code
  warning = FALSE,       # don't show warnings
  message = FALSE,       # don't show messages (less serious warnings)
  cache = FALSE,         # set to TRUE to save results from last compilation
  # pdf.options(family="GB1") #show chinese in plots
  fig.align = 'center',  # center figures
  # out.width = '75%',    # Set figure size to 75% of textline width
  # fig.asp = .75         # width-height ratio 
  fig.pos = 'p',         # Places figures on their own pages
  out.width = '100%', 
  dpi=300
)
# I usually load my libraries up front to keep things organized
library(bookdown)
library(tidyverse)
#library(kableExtra)
#library(ggthemes)
#library(ggExtra)
set.seed(2018)         # make random results reproducible
```


# General form of a research paper

> Writing is easy. All you do is stare at a blank sheet of paper until drops of blood form on your forehead. 
> --- Gene Fowler

Written and oral communications skills are probably the most universal qualities sought by graduate and professional schools as well as by employers. You alone are responsible for developing such skills to a high level.

An objective of organizing a research paper is to allow people to read your work selectively. When I research a topic, I may be interested in just the methods, a specific result, the interpretation, or perhaps I just want to see a summary of the paper to determine if it is relevant to my study. To this end, many journals require the following sections, submitted in the order listed, each section to start on a new page. There are variations of course. Some journals call for a combined results and discussion, for example, or include materials and methods after the body of the paper. The well known journal *Science* does away with separate sections altogether, except for the abstract.

Your papers are to adhere to the form and style required for the *Journal of Biological Chemistry*, requirements that are shared by many journals in the life sciences.

This guideline is from *Experimental Biosciences Resources, Rice University*    
<http://www.ruf.rice.edu/~bioslabs/tools/report/reportform.html#form>.

Reading material: McMillan, VE. "Writing Papers in the Biological Sciences, six Ed." [@mcmillan_writing_2016].

# General style

Specific editorial requirements for submission of a manuscript will always supercede instructions in these general guidelines.

## To make a paper readable

- Print or type using a 12 point standard font, such as Times, Geneva, Bookman, Helvetica, etc.
- Text should be double spaced on 8 1/2" $\times$ 11" paper with 1 inch margins, single sided
Number pages consecutively
- Start each new section on a new page
- Adhere to recommended page limits

## Mistakes to avoid

- Placing a heading at the bottom of a page with the following text on the next page (insert a page break!)
- Dividing a table or figure - confine each figure/table to a single page
- Submitting a paper with pages out of order

## In all sections of your paper

- Use normal prose including articles ("a", "the," etc.)
- Stay focused on the research topic of the paper
- Use paragraphs to separate each important point (except for the abstract)
- Indent the first line of each paragraph
- Present your points in logical order
- Use present tense to report well accepted facts - for example, 'the grass is green' 
- Use past tense to describe specific results - for example, 'When weed killer was applied, the grass was brown' 
- Avoid informal wording, don't address the reader directly, and don't use jargon, slang terms, or superlatives 
- Avoid use of superfluous pictures - include only those figures necessary to presenting results


# Title Page

Select an informative title. Include the name(s) and address(es) of all authors, and date submitted.

# Abstract

The summary should be two hundred words or less.

## General intent

An abstract is a concise single paragraph summary of completed work or work in progress. In a minute or less a reader can learn the rationale behind the study, general approach to the problem, pertinent results, and important conclusions or new questions. 

## Writing an abstract

Write your summary after the rest of the paper is completed. After all, how can you summarize something that is not yet written? Economy of words is important throughout any paper, but especially in an abstract. However, use complete sentences and do not sacrifice readability for brevity. You can keep it concise by wording sentences so that they serve more than one purpose. For example, "In order to learn the role of protein synthesis in early development of the sea urchin, newly fertilized embryos were pulse-labeled with tritiated leucine, to provide a time course of changes in synthetic rate, as measured by total counts per minute (cpm)." This sentence provides the overall question, methods, and type of analysis, all in one sentence. The writer can now go directly to summarizing the results.

Summarize the study, including the following elements in any abstract. Try to keep the first two items to no more than one sentence each.

- Purpose of the study - hypothesis, overall question, objective
- Model organism or system and brief description of the experiment
- Results, including specific data - if the results are quantitative in nature, report quantitative data; results of any statistical analysis shoud be reported
- Important conclusions or questions that follow from the experiment(s)

## Style

- Single paragraph, and concise
- As a summary of work done, it is always written in past tense
- An abstract should stand on its own, and not refer to any other part of the paper such as a figure or table
- Focus on summarizing results - limit background information to a sentence or two, if absolutely necessary
- What you report in an abstract must be consistent with what you reported in the paper
- Corrrect spelling, clarity of sentences and phrases, and proper reporting of quantities (proper units, significant figures) are just as important in an abstract as they are anywhere else

# Introduction

Your introductions should not exceed two pages (double spaced, typed). 

## General intent

The purpose of an introduction is to aquaint the reader with the rationale behind the work, with the intention of defending it. It places your work in a theoretical context, and enables the reader to understand and appreciate your objectives.

## Writing an introduction

The abstract is the only text in a research paper to be written without using paragraphs in order to separate major points. Approaches vary widely, however for our studies the following approach can produce an effective introduction.

- Describe the importance (significance) of the study - why was this worth doing in the first place? Provide a broad context.
- Defend the model - why did you use this particular organism or system? What are its advantages? You might comment on its suitability from a theoretical point of view as well as indicate practical reasons for using it.
- Provide a rationale. State your specific hypothesis(es) or objective(s), and describe the reasoning that led you to select them.
- Very briefy describe the experimental design and how it accomplished the stated objectives.

## Style

- Use past tense except when referring to established facts. After all, the paper will be submitted after all of the work is completed. 
- Organize your ideas, making one major point with each paragraph. If you make the four points listed above, you will need a minimum of four paragraphs.
- Present background information only as needed in order support a position. The reader does not want to read everything you know about a subject.
- State the hypothesis/objective precisely - do not oversimplify.
- As always, pay attention to spelling, clarity and appropriateness of sentences and phrases.


# Materials and Methods 

There is no specific page limit, but a key concept is to keep this section as concise as you possibly can. People will want to read this material selectively. The reader may only be interested in one formula or part of a procedure. Materials and methods may be reported under separate subheadings within this section or can be incorporated together. 

## General intent

This should be the easiest section to write, but many students misunderstand the purpose. The objective is to document all specialized materials and general procedures, so that another individual may use some or all of the methods in another study or judge the scientific merit of your work. It is not to be a step by step description of everything you did, nor is a methods section a set of instructions. In particular, it is not supposed to tell a story. By the way, your notebook should contain all of the information that you need for this section.

## Writing a materials and methods section

## Materials

- Describe materials separately only if the study is so complicated that it saves space this way.
- Include specialized chemicals, biological materials, and any equipment or supplies that are not commonly found in laboratories. 
- Do not include commonly found supplies such as test tubes, pipet tips, beakers, etc., or standard lab equipment such as centrifuges, spectrophotometers, pipettors, etc.
- If use of a specific type of equipment, a specific enzyme, or a culture from a particular supplier is critical to the success of the experiment, then it and the source should be singled out, otherwise no.
- Materials may be reported in a separate paragraph or else they may be identified along with your procedures.
- In biosciences we frequently work with solutions - refer to them by name and describe completely, including concentrations of all reagents, and pH of aqueous solutions, solvent if non-aqueous.

## Methods

- Report the methodology (not details of each procedure that employed the same methodology)
- Describe the mehodology completely, including such specifics as temperatures, incubation times, etc.
- To be concise, present methods under headings devoted to specific procedures or groups of procedures
- Generalize - report how procedures were done, not how they were specifically performed on a particular day. For example, report "samples were diluted to a final concentration of 2 mg/ml protein;" don't report that "135 microliters of sample one was diluted with 330 microliters of buffer to make the protein concentration 2 mg/ml." Always think about what would be relevant to an investigator at another institution, working on his/her own project.
- If well documented procedures were used, report the procedure by name, perhaps with reference, and that's all. For example, the Bradford assay is well known. You need not report the procedure in full  \- just that you used a Bradford assay to estimate protein concentration, and identify what you used as a standard. The same is true for the SDS-PAGE method, and many other well known procedures in biology and biochemistry.

## Style 

- It is awkward or impossible to use active voice when documenting methods without using first person, which would focus the reader's attention on the investigator rather than the work. Therefore when writing up the methods most authors use third person passive voice.
- Use normal prose in this and in every other section of the paper – avoid informal lists, and use complete sentences.

## What to avoid

- Materials and methods are not a set of instructions.
- Omit all explanatory information and background - save it for the discussion.
- Omit information that is irrelevant to a third party, such as what color ice bucket you used, or which individual logged in the data.


# Results

The page length of this section is set by the amount and types of data to be reported. Continue to be concise, using figures and tables, if appropriate, to present results most effectively. See recommendations for content, below. 

## General intent

The purpose of a results section is to present and illustrate your findings. Make this section a completely objective report of the results, and save all interpretation for the discussion.

## Writing a results section

IMPORTANT: You must clearly distinguish material that would normally be included in a research article from any raw data or other appendix material that would not be published. In fact, such material should not be submitted at all unless requested by the instructor.

## Content

- Summarize your findings in text and illustrate them, if appropriate, with figures and tables. 
In text, describe each of your results, pointing the reader to observations that are most relevant.
- Provide a context, such as by describing the question that was addressed by making a particular observation.
- Describe results of control experiments and include observations that are not presented in a formal figure or table, if appropriate.
- Analyze your data, then prepare the analyzed (converted) data in the form of a figure (graph), table, or in text form.

## What to avoid

- Do not discuss or interpret your results, report background information, or attempt to explain anything.
- Never include raw data or intermediate calculations in a research paper.
- Do not present the same data more than once.
- Text should complement any figures or tables, not repeat the same information. 
- Please do not confuse figures with tables - there is a difference.

## Style

- As always, use past tense when you refer to your results, and put everything in a logical order.
- In text, refer to each figure as "figure 1," "figure 2," etc. ; number your tables as well (see the reference text for details)
- Place figures and tables, properly numbered, in order at the end of the report (clearly distinguish them from any other material such as raw data, standard curves, etc.)
- If you prefer, you may place your figures and tables appropriately within the text of your results section.

## Figures and tables

- Either place figures and tables within the text of the result, or include them in the back of the report (following Literature Cited) - do one or the other
- If you place figures and tables at the end of the report, make sure they are clearly distinguished from any attached appendix materials, such as raw data
- Regardless of placement, each figure must be numbered consecutively and complete with caption (caption goes under the figure) 
- Regardless of placement, each table must be titled, numbered consecutively and complete with heading (title with description goes above the table) 
- Each figure and table must be sufficiently complete that it could stand on its own, separate from text

# Discussion

Journal guidelines vary. Space is so valuable in the *Journal of Biological Chemistry*, that authors are asked to restrict discussions to four pages or less, double spaced, typed. That works out to one printed page. While you are learning to write effectively, the limit will be extended to five typed pages. If you practice economy of words, that should be plenty of space within which to say all that you need to say. 

## General intent

The objective here is to provide an interpretation of your results and support for all of your conclusions, using evidence from your experiment and generally accepted knowledge, if appropriate. The significance of findings should be clearly described. 

## Writing a discussion

Interpret your data in the discussion *in appropriate depth*. This means that when you explain a phenomenon you must describe mechanisms that may account for the observation. If your results differ from your expectations, explain why that may have happened. If your results agree, then describe the theory that the evidence supported. It is never appropriate to simply state that the data agreed with expectations, and let it drop at that.

- Decide if each hypothesis is supported, rejected, or if you cannot make a decision with confidence. - Do not simply dismiss a study or part of a study as "inconclusive."
- Research papers are not accepted if the work is incomplete. Draw what conclusions you can based upon the results that you have, and treat the study as a finished work 
- You may suggest future directions, such as how the experiment might be modified to accomplish another objective.
- Explain all of your observations as much as possible, *focusing on mechanisms*.
- Decide if the experimental design adequately addressed the hypothesis, and whether or not it was properly controlled.
- Try to offer alternative explanations if reasonable alternatives exist.
- One experiment will not answer an overall question, so keeping the big picture in mind, where do you go next? The best studies open up new avenues of research. What questions remain?
- Recommendations for specific papers will provide additional suggestions.

## Style 

- When you refer to information, distinguish data generated by your own studies from published information or from information obtained from other students (verb tense is an important tool for accomplishing that purpose). 
- Refer to work done by specific individuals (including yourself) in past tense. 
- Refer to generally accepted facts and principles in present tense. For example, "Doofus, in a 1989 survey, *found* that anemia in basset hounds *was correlated* with advanced age. Anemia *is* a condition in which there *is* insufficient hemoglobin in the blood."

The biggest mistake that students make in discussions is to present a superficial interpretation that more or less re-states the results. It is necessary to suggest *why* results came out as they did, focusing on the mechanisms behind the observations.

# Literature Cited

Please note that in the introductory laboratory course, you will not be required to properly document sources of all of your information. One reason is that your major source of information is this website, and websites are inappropriate as primary sources. Second, it is problematic to provide a hundred students with equal access to potential reference materials. You may nevertheless find outside sources, and you should cite any articles that the instructor provides or that you find for yourself.

List all literature cited in your paper, in alphabetical order, by first author. In a proper research paper, only primary literature is used (original research articles authored by the original investigators). Be cautious about using web sites as references - anyone can put just about anything on a web site, and you have no sure way of knowing if it is truth or fiction. If you are citing an on line journal, use the journal citation (name, volume, year, page numbers). Some of your papers may not require references, and if that is the case simply state that "no references were consulted." 

# Selekted Writing Rules

1. Verbs HAS to agree with their subjects. 
2. Prepositions are not words to end sentences with. 
3. And don't start a sentence with a conjunction.
4. It is wrong to ever split an infinitive. 
5. Avoid cliches like the plague. (They're old hat) 
6. Also, always avoid annoying alliteration. 
7. Be more or less specific. 
8. Parenthetical remarks (however relevant) are (usually) unnecessary.
9. Also too, never, ever use repetitive redundancies.
10. No sentence fragments. 
11. Contractions aren't necessary and shouldn't be used. 
12. Foreign words and phrases are not apropos. 
13. Do not be redundant; do not use more words than necessary; it's highly superfluous. 
14. One should NEVER generalize. 
15. Comparisons are as bad as cliches. 
16. Eschew ampersands & abbreviations, etc. 
17. One‐word sentences? Eliminate.
18. Analogies in writing are like feathers on a snake. 
19. The passive voice is to be ignored. 
20. Eliminate commas, that are, not necessary. Parenthetical words however should be enclosed in  commas.
21. Never use a big word when a diminutive one would suffice. 
22. Use words correctly, irregardless of how others use them. 
23. Understatement is always the absolute best way to put forth earth‐shaking ideas.
24. Eliminate quotations. As Ralph Waldo Emerson said, "I hate quotations. Tell me what you know." 
25. If you've heard it once, you've heard it a thousand times: Resist hyperbole; not one writer in a million  can use it correctly. 
26. Puns are for children, not groan readers. 
27. Go around the barn at high noon to avoid colloquialisms.
28. Even IF a mixed metaphor sings, it should be derailed.
29. Who needs rhetorical questions?
30. Exaggeration is a billion times worse than understatement.
31. Proofread carefully to see if you any words out.


# Appendix{-}

Use `\usepackage{endfloat}` put all figures to the end.

Figure \@ref(fig:points): by ggplot2.


```{r points, warning = 'FALSE', fig.cap='ggplot2 with facet and variable point size'}
library(ggplot2)
p <- ggplot(data = mpg, mapping = aes(x = cty, y = hwy))
p + geom_point(aes(colour = class, size = displ), alpha = 0.5) +
  stat_smooth() + scale_size_continuous(range = c(4, 10)) +
  facet_wrap(~ year, ncol = 1)
```

Figure \@ref(fig:multiplot): Two plots placed side by side.

```{r multiplot, out.width='50%', fig.show='hold', fig.cap='Two plots placed side by side.'}
par(mar = c(4, 4, .1, .1))
plot(pressure, pch = 19, type = 'b')
plot(cars, pch = 19)
```

# References{-}
